---
title: "July 2020 Puzzle"
author: "Calum Davey"
date: "25/07/2020"
output: 
  html_document:
    theme: readable
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Puzzle  

I have a dataset with 5 different x values and the lower limit of the difference from x = 8. (limit always = 0 when x = 8)  

I would like a function to get the lowest x where limit < 0.05 and all of the previous limits in the sequence to 8 are also < 0.05.   

I don't mind if the function returns a single value or a dataset with one row. Example datasets and expected answers are below.   

## Datasets and expected answers

```{r ProblemDatasets}
# f(a) = 4
a <- data.frame(x = c(8, 7, 6, 5, 4), 
                limit = c(0, 0.01, 0.01, 0.01, 0.01))

# f(b) = 6
b <- data.frame(x = c(8, 7, 6, 5, 4), 
      limit = c(0, 0.01, 0.01, 0.06, 0.06))

# f(c) = 7 (as 6 is outside of limit)
c <- data.frame(x = c(8, 7, 6, 5, 4), 
                limit = c(0, 0.01, 0.06, 0.01, 0.06))

# f(d) = 7 (as 6 is outside of limit)
d <- data.frame(x = c(8, 7, 6, 5, 4), 
                limit = c(0, 0.01, 0.06, 0.01, 0.01))
```

## Solutions
### Calum 

```{r Calum}
f <- function(Y, X, threshold, d){
  p <- suppressWarnings(min(which((d[,Y] - threshold) > 0)) - 1)
  if(p == Inf){p <- nrow(d)}
  d[p, X]  
}
```

### Antonio 

```{r Antonio}
# FUNCTION fweird WITH ARGUMENTS:
# - x: VALUES
# - limit: CORRESPONDING DECIMALS
# - thr: THRESHOLD FOR DECIMAL
# - res: RETURN VALUE OR POSITION
# NB: SET TO NA IF NONE IS ELIGIBLE
fweird <- function(x, limit, thr=0.05, res="val") {
  # CHECK CONSISTENCY OF ARGUMENTS AND THEIR VALUE
  if(length(x)!=length(limit)) 
    error("'x' and 'limit' must have the same length")
  res <- match.arg(res, c("val","ind"))
  
  # IDENTIFY WHICH DECIMALS ARE ABOVE THE THRESHOLD
  not <- which(limit>=0.05)
  
  # IDENTIFY THE INDEX
  ind <- ifelse(length(not), ifelse(not[1]==1, NA, not[1]-1), length(x))
  
  # RETURN NA IF NONE
  if(is.na(ind)) return(NA)
  
  # RETURN EITHER THE VALUE OR THE INDEX
  ifelse(res=="val", x[ind], ind)
}
```

### Darrren 

```{r Darren}
#D Scott function

min_fun = function(data, alpha){
  #================================================
  #Purpose#
  #(Find the lowest x where limit < 0.05 and all of 
  # the previous limits in the sequence.)
  #------------------------------------------------
  #Parameters#
  #data = Data set, 2 columns X and limit
  #alpha = Threshold  
  #------------------------------------------------  
  #Output#  
  #Min of x up to limit
  #================================================
  
  indx = seq(1, length(data$x))
  indco = indx[data$limit >= alpha][1]
  
  if(is.na(indco)){
    
    max = length(data$x)
    print("Threshold not in x")
    se = seq(1, max)
    min_x = min(data$x[se])
    
  }else if(indco > 1){
    max = indco
    se = seq(1, max - 1)
    min_x = min(data$x[se])
    
  }else{
    min_x = NA
    print("Threshold is the first value")
  }
  
  return(min_x)
}
```

### Clemence 

```{r Clemence}
#Function: first check if any limit is larger than the threshold. 
#If not, the result is the last 
#value of the vector x. Otherwise, look at the first time 
#limit>threshold and get the x value from previous row)

puzzle<-function(df){
  result<-ifelse(sum(df$limit>threshold)==0, tail(df$x, n=1),
       df$x[min(which(df$limit>threshold))-1])
  return(result)
}
```

## Alex 

```{r Alex}
puzzlefunction <- function(data,threshold=0.05){
  index <- sum(1-sign(cumsum(data$limit >= 0.05)))  
  return(data$x[index])
}
```

### Speedtests

```{r SpeedTests}
data <- b
alpha <- threshold <- 0.05

# Benchmarking the solution 
  # install.packages('microbenchmark')
  library(microbenchmark)
  m <- microbenchmark(
  
  # Calum
    f(Y = 'limit', X = 'x', threshold = 0.05, d = data),
    
  # Antonio 
    fweird(x = data$x, limit = data$limit),
  
  # Darren 
    min_fun(data, alpha),
  
  # Clemence 
    puzzle(data),
  
  # Alex 
    puzzlefunction(data, threshold),
  
  times = 1000
) 
  
barplot(summary(m)$median,
        names = c( 'Calum', 'Antonio', 'Darren', 'Clemence', 'Alex'),
        ylab = 'Median time in ns')  

```

